---
pdf_document: null
author: "Kieran O'Neill"
highlight: tango
output:
  pdf_document:
    keep_tex: yes
keep_md: yes
html_document: null
title: "A (Very) Short Introduction to R for Wet Lab Scientists"
fig_caption: no
toc: yes
toc_depth: 2
---
	
# What is R, and Why Should I Use It?
	
## The Basics

R is a versatile, open source programming language that was specifically designed for data analysis. As such R is extremely useful both for statistics and data science. Inspired by the programming language S.  

* Open source software under GPL.  
* Superior (if not just comparable) to commercial alternatives. R has over 5,000 user contributed packages at this time. It's widely used both in academia and industry.  
* Available on all platforms. 
* Large and growing community of peers.
* Bioconductor: largest (and arguably the best) free collection of software for biological data analysis anywhere.

## Why Not Just Use Excel, FlowJo, GraphPad, etc?

1. Reproducibility
	* Its really important that you know what you did. 
	* More journals/grants/etc. are also requiring this. 
	* The best way to know what you did is to provide all the code.
	* GUI software makes this difficult
	* If you keep a lab notebook, why not do the same thing with your analysis?
	
2. Flexibility, capabilities and pretty pictures
	* R can handle much larger data sets, much faster, and much more easily than Excel.
	* Huge range of statistical tests, biological data types, etc.
	* Plotting in R is far more sophisticated than any available GUI.


## Proof of What I Mean By Pretty Pictures:

![[Gerstung et al (2015) Nature Communications (CC-BY)](http://www.nature.com/ncomms/2015/150109/ncomms6901/full/ncomms6901.html#supplementary-information) ](images/ncomms6901-f1.jpg)



# RStudio

## Set up a new project

- Click 'file', then 'New project'
- Click 'New directory'm then 'Empy Project', then pick a directory
- With the project set up, click 'file', then 'new' (or ctrl+shift+n)
- Click 'File', 'Save' (ctrl+s)
- Save the file as something meaningful, like `lecture1_examples.R`

Note: for Mac users, where I say 'ctrl', use your weird Mac control key instead.

## Quick overview of RStudio

-Figure of interface

## Working between the script and console {.fragile}

*Type the following into the console, and press enter:

```{r}
print("Hello")
```

* Now type it into the file window, and with the cursor on that line, press ctrl+enter
* Messing around in the console is fun.
* But it's better to keep your work in a file which you save often.


# R Basics

## Objects

You can assign values to objects:

```{r}
some_number <- 5
some_number + 3
some_other_number <- some_number ^ 3
some_other_number
```

Take a look in your environment window in RStudio. You can also see what objects are defined using the `ls()` command:

```{r}
ls()
```

## Basic Data Types

You can find out the type of an object using `typeof()`:

```{r}
typeof(some_number)
```

```{r}
some_text <- "5" 
typeof(some_text)
```

## Numeric vs Character

```{r}
some_number + 5 
```

```{r}
some_text + 5
```

```{r}
as.numeric(some_text) + 5
```

## Vectors

Vectors are one-dimensional objects. You can create them with the `c()` function:

```{r}
my_vector <- c(1,3,5,6,7,8)
```

You can apply operations to a whole vector.

```{r}
my_vector^3
```

You can join vectors with `c()`:

```{r}
c(my_vector, my_vector^3, my_vector^4)
```

## Data frames

Data frames are special objects in R for storing mixed data. You can think of them as something like a sheet in Excel. We're going to load in one of R's default data sets, a series



# Installing Packages

## CRAN

## Bioconductor

# Loading in Files

## The Iris Data Set

## Loading in CSV

## Loading in from Excel ?



# Basic Statistical Tests

## Fisher's T Test

## Examining the Results

# Basic Plots

## Scatter Plot

## Box Plots

- Don't use dynamite plots!!

## Beeswarm Plots

## Other Plotting Packages

# Credits

## Course Developers

- Kieran O'Neill
- Eva Yap
- Alice Zhu (for next session)

Much material was reused from [Software Carpentry's Bootcamp workshops](https://github.com/swcarpentry/bc/tree/master/novice/r) and from [Andy Teucher's short R course](https://github.com/ateucher/rcourse_site), both under the terms of the [Creative Commons Attribution License](http://creativecommons.org/licenses/by/3.0/). 
